### 2022-11-30 paper1.tex
ie and eg

Let $\theta$ denote the effect of a treatment on an outcome of interest. Let
$\that_o$ and $\that_r$ denote its maximum likelihood estimates obtained from an
original (subscript $o$) and from a replication study (subscript $r$),
respectively. Let the corresponding standard errors be denoted by $\sigma_o$ and
$\sigma_r$, the $z$-values by $z_o = \that_o/\sigma_o$ and
$z_r = \that_r/\sigma_r$, and define the variance ratio as
$c = \sigma_o^2/\sigma_r^2$ and the relative effect estimate as
$d = \that_r/\that_o = z_r/(z_o\sqrt{c})$. For many effect size types the
variances are inversely proportional to the sample size, \ie
$\sigma^2_{o} = \kappa/n_{o}$ and $\sigma^2_{r} = \kappa/n_{r}$ for some unit
variance $\kappa$. The variance ratio is then the ratio of the replication to
the original sample size $c = n_{r}/n_{o}$.

We adopt a meta-analytic framework and consider the effect estimates as the
data, rather than their underlying samples, and assume that
$\that_k \given \theta \sim \Nor(\theta, \sigma^2_k)$ for $k \in \{o, r\},$ \ie
normality of the effect estimates around $\theta$, with known variances equal to
their squared standard errors. For studies with reasonable sample size, this
framework usually provides a good approximation for a wide range of (suitably
transformed) effect size types \citep[Chapter 2.4]{Spiegelhalter2004}. For
example, means and mean differences (no transformation), odds ratios, hazard
ratios, risk ratios (logarithmic transformation), or correlation coefficients
(``Fisher-$z$'' transformation). We refer to the literature of meta-analysis for
details about transformations of effect sizes \citep[\eg][Chapter
11.6]{Cooper2019}. The normal model in combination with conjugate priors enables
derivation of closed-form expressions in many cases, which allows us to easily
study limiting behaviour and facilitates interpretability. In Section
\ref{sec1:tdist}, we will present relaxations of the normality assumption, which
can lead to more accurate inferences when studies have small sample sizes and/or
show extreme results.

\section{Reverse-Bayes assessment of replication success with Bayes factors}
\label{sec1:methods}
% ``judgements are made concerning the values that the final
% probabilities would have after some imaginary results are obtained; then Bayes'
% theorem is used backwards to estimate what the consistent initial probabilities
% must be assumed to be; and lastly Bayes's theorem is used (forwards) to obtain
% estimates of the final probabilities in the light of the actual observations.
% This device takes \emph{seriously} the view that a theory of probability or of
% rationality is no more than a theory of consistency'' (p. 126).
The idea of reversing Bayes' theorem was first proposed by \citet{Good1950}. He
acknowledged that in many situations there is no obvious choice for the prior
distributions involved in Bayesian analyses. On the other hand, we are often
more certain which posterior inferences would convince us regarding the
credibility of a hypothesis. For this reason, Good inverted Bayes' theorem and
derived priors, which combined with the observed data, would lead to posterior
inferences that were specified beforehand (\eg the data favour one hypothesis
over another). His reverse-Bayes inference then centred around the question
whether the resulting prior is plausible, and if so, this would legitimise the
posterior inference. See Figure~\ref{fig1:RBschmema} for a graphical illustration
of this process.

@@

Good argued that philosophically there is nothing wrong with inferences
resulting from backwards use of Bayes' theorem, since the theorem merely
constrains prior and posterior to be consistent with the laws of probability
(regardless of their conventional names suggesting a particular temporal
ordering). Despite his advocacy, the reverse-Bayes idea remained largely
unexplored until \citet{Matthews2001b} introduced the \emph{Analysis of
Credibility}, which in turn led to new developments in reverse-Bayes methodology
(see \citet{Held2021b} for a recent review). Most of these approaches use the
reversal of Bayes' theorem in order to challenge or substantiate the credibility
of scientific claims. Usually, a posterior inference corresponding to
(non-)credibility of a claim is specified, and the associated prior is then
derived from the data. Inference is subsequently carried out based on this
reverse-Bayes prior, \eg the interest is often to check whether the prior is
plausible in light of external evidence, an obvious candidate being data from a
replication study. This can be done, for example, using methods to assess
prior-data conflict \citep{Box1980, Evans2006}.

@@

    \item \textbf{Replication study}:
          For the replication effect estimate
$\that_r \given \theta \sim \Nor(\theta, \sigma^2_r)$ compute the Bayes factor
contrasting the sceptic $\HS \colon \theta \sim \Nor(0, \gy \cdot \sigma^2_o)$
to an advocate $\HA \colon \theta \sim \Nor(\that_o, \sigma^2_o)$. Declare
\emph{replication success} at level $\gamma$ if
          $$\BFcol{\mathrm{S}}{\mathrm{A}}(\that_r; \gy) \leq \gamma,$$
          \ie the data favour the advocate over the sceptic at a higher level
than the sceptic's initial objection.

@@

The Bayes factor from equation ~\eqref{eq1:BFo} is shown in Figure ~\ref{fig1:bfo}
as a function of $g$ and for different original $z$-values $z_o$. For fixed
$z_o$, it is well known that this Bayes factor is bounded from below by
\begin{align}
  \label{eq1:minBFo}
  \minBF_{o} =
  \begin{cases}
    |z_o| \cdot  \exp(-z_o^2/2) \cdot \sqrt{e} & \text{for} ~ |z_o| > 1 \\
    1 & \text{for} ~ |z_o| \leq 1
  \end{cases}
\end{align}
which is reached at $\gminBFo = \max\{0, z^2_o - 1\}$ \citep{Edwards1963}.
Further increasing the relative sceptical prior variance increases
\eqref{eq1:BFo} indefinitely because of the Jeffreys-Lindley paradox, \ie
$\BFcol{0}{\mathrm{S}}(\hat{\theta}_o; g) \to \infty$ for $g \to \infty$
\citep[][Section 6.1.4]{Bernardo2000}. Hence, for a relative sceptical prior
variance $g \in [0, \gminBFo]$, the resulting Bayes factor will be
$\BFcol{0}{\mathrm{S}}(\hat{\theta}_o; g) \in [\minBF_o, 1]$.

@@

From Figure \ref{fig1:bfo} we can see that the more compelling the original data
(\ie the larger $|z_o|$), the smaller the sufficiently sceptical relative prior
variance $\gy$ needs to be in order to make the result no longer convincing at
level $\gamma$. In the most extreme case, when $|z_{o}| \to \infty$ and $\gamma$
remains fixed, the sufficiently sceptical prior variance will converge to
zero (Appendix \ref{appendix:lambertW}). On the other hand, if $|z_o|$ is not
sufficiently large, $\BFcol{0}{\mathrm{S}}(\hat{\theta}_o; g)$ will either be
always increasing in $g$ (if $|z_o| \leq 1$) or it will reach a minimum above
the chosen level $\gamma$. In both cases the sufficiently sceptical relative
prior variance $\gy$ is not defined since there is no need to challenge an
already unconvincing result.

A side note on the Jeffreys-Lindley paradox is worth being mentioned: If a
$\gy < \gminBFo$ exists, there exists also a $\gy^\prime > \gminBFo$ as the
Bayes factor monotonically increases in $g > \gminBFo$ and therefore must
intersect a second time with $\gamma$, due to the paradox. This means that the
more compelling the original result, the larger $\gy^\prime$ needs to be chosen,
such that the result becomes no longer convincing at level $\gamma$. However,
priors which become increasingly diffuse do not represent increasing scepticism
but rather increasing ignorance. Using \eqref{ggamma} therefore avoids this
manifestation of the Jeffreys-Lindley paradox, since it determines sceptical
priors only from the class of priors that become increasingly concentrated for
increasing evidence (\ie priors with $\gy \leq \gminBFo$). In principle, the
solution $\gy^\prime > \gminBFo $ could also be computed by replacing the
$\lw{-1}$ branch of the Lambert $W$ function in \eqref{ggamma} with the $\lw{0}$
branch, but this will not be of interest to us.

@@

For example, if we observe $z_o = 3$ (equivalent to minimum Bayes factor
$\minBF_{o} = 1/18$) and choose a level $\gamma = 1/10$
the sufficiently sceptical relative prior variance \eqref{ggamma} is
$\gy = 1.6$. If a replication is conducted with the
same precision ($c = 1$) and we observe $z_r = 2.5$ (equivalent to
minimum Bayes factor $\minBF_{r} = 1/5.5$ and relative effect
estimate $d = z_{r}/(z_{o}\sqrt{c}) = 0.83$), using equation
\eqref{eq1:BFr} this would lead to $\BFcol{\mathrm{S}}{\mathrm{A}}(\hat{\theta}_r;
1.6) = 1/3.5$, which means
that the replication was not successful at level $\gamma = 1/10$.
However, if we had chosen a less stringent level, \eg$\gamma =
1/3$, the replication would have been considered successful
since then $\gy = 0.4$ and
$\BFcol{\mathrm{S}}{\mathrm{A}}(\hat{\theta}_r; 0.4)
= 1/7.4$.

@@

\subsubsection{The two-trials rule}
Replication success is most commonly declared when both original and replication
study provide compelling evidence against a null effect. This approach is also
known as the \emph{two-trials rule} in drug development and usually a
requirement for drug approval \citep[Section 9.4]{Kay2015}. Most replication
projects report $p$-values associated with the effect estimates as measures of
evidence against the null, but also default Bayes factors have been used
\citep[see \eg the Bayesian supplement of][]{Camerer2018}. To compare the
two-trials rule with methods based on Bayes factors we will study the two-trials
rule based on the minimum Bayes factor from~\eqref{eq1:minBFo}, \ie replication
success at level $\gamma$ is established when both
$\minBF_{k} % = \abs{z_{k}}\exp(-z_{k}^{2}/2)\sqrt{e}
\leq \gamma$ for $k \in \{o, r\}$, as well as
$\sign(\hat{\theta}_o) = \sign(\hat{\theta}_r)$. This approach has a one-to-one
correspondence to the usual version of two-trials rule as minimum Bayes factors
and $p$-values both only depend on the $z$-values of original and replication
study.

The two-trials rule guarantees that both studies provide compelling evidence
against the null. Similarly, the sceptical Bayes factor requires the original
study to be compelling on its own since it can never be smaller than
$\minBF_{o}$. However, one can easily construct examples where the sceptical
Bayes factor is smaller than the minimum Bayes factor from the replication study
(\eg when $\minBF_{o} = 1/2$,
$\minBF_{r} = 1/1.5$, and $c = 1$
we obtain $\BFs = 1/1.9$). So for the same level of
replication success $\gamma$ the two-trials may not flag replication success
whereas the sceptical Bayes factor would.

@@

Similarly, as with the sceptical Bayes factor, the $Q$-statistic
from~\eqref{eq1:Qstat} appears in~\eqref{eq1:BFrep} and acts as a penalty term,
\ie larger values of $Q$ lower the degree of replication success. However, in
contrast to the sceptical Bayes factor, the replication Bayes factor is not
limited by the evidence from the original study because $\BFr \downarrow 0$ as
$c \to \infty$ provided $z_{o} \neq 0$ and $d \neq 0$. Moreover, we have that
\begin{align*}
  1 \geq \BFs \geq \BFcol{\mathrm{S}}{\mathrm{A}}(\hat{\theta}_r; g_{\scriptscriptstyle \BFs})
  = \BFcol{\mathrm{S}}{0}(\hat{\theta}_r; g_{\scriptscriptstyle \BFs}) \cdot \underbrace{\BFcol{0}{\mathrm{A}}(\hat{\theta}_r)}_{= \BFr}.
\end{align*}
So the sceptical Bayes factor is larger than the replication Bayes factor if the
replication data favour the sceptical prior
$\HS \colon \theta \sim \Nor(0, g_{\scriptscriptstyle \BFs} \cdot \sigma^{2}_{o})$
over the null hypothesis. They can only coincide when $\BFs = 1$ since then
$g_{\scriptscriptstyle \BFs} = 0$.

@@

\subsubsection{The sceptical \textit{p}-value}
Of particular interest is the relationship between the sceptical Bayes factor
and the sceptical $p$-value \citep{Held2020}, as it is the outcome of a similar
reverse-Bayes procedure. One also considers a sceptical prior for the effect
size $\theta \sim \Nor(0, \tau^2)$, the sufficiently sceptical prior variance at
level $\alpha$ is then defined as $\tau^2 = \tau^2_\alpha$ such that the
$(1 - \alpha)$ credible interval for $\theta$ based on the posterior
$\theta \given \that_o, \tau^2_\alpha$ does not include zero. Replication
success is declared if the tail probability of the replication effect estimate
under its prior predictive distribution
$\that_r \given \tau^2_\alpha \sim \Nor(0, \tau^2_\alpha + \sigma^2_r)$ is
smaller than $\alpha$. The smallest level $\alpha$ where replication success can
be established defines the sceptical $p$-value. In contrast to the sceptical
Bayes factor, the sceptical $p$-value always exists and there are closed form
expressions to compute it for all values of $c$, \ie
$\ps = 1 - \Phi(z_{\mathrm{S}})$ with
\begin{equation*}
  z^2_{\mathrm{S}} =
  \begin{cases}
    z^2_H/2 & \text{for}~c=1 \\ \frac{1}{c - 1}\left\{\sqrt{z^2_A\left[z^2_A +
          (c - 1)z^2_H\right]} - z^2_A\right\} & \text{for}~c \neq 1
  \end{cases}
  \label{eq1:z2s}
\end{equation*}
where $z^2_H = 2/(1/z_o^2 + 1/z_r^2)$ the harmonic mean,
$z^2_A = (z_o^2 + z_r^2)/2$ the arithmetic mean of the squared $z$-statistics,
and provided that $\sign(\that_o) = \sign(\that_r)$ (otherwise
$\ps = \Phi(z_{\mathrm{S}})$).

@@

In practice, the replication paradox is hardly an issue, since replications
rarely show such contradictory results, \eg to achieve replication success at
level $\gamma = 1/3$ with
$\minBF_{o} = 1/10$ and $c = 1$, the
relative effect estimate needs to be $d < -7.09$ for the
paradox to appear with the sceptical Bayes factor. The replication Bayes factor
is more prone to the paradox because its point-null hypothesis fails more
strongly to predict estimates in opposite direction, \eg for the same numbers as
before it requires $d < -2.66$.

In both cases the paradox can be overcome by truncating the advocacy prior $\HA$
such that only effects in the same direction as the original estimate $\that_o$
have non-zero probability, \ie for positive $\that_{o}$ consider
$\HAp \colon \theta \sim \Nor(\that_o, \sigma^2_o) \, \mathbb{1}_{(0, \infty)}(\theta)$,
where $\mathbb{1}_{B}(x)$ is the indicator function of the set $B$. The Bayes
factor contrasting $\HS$ to $\HAp$ turns out to be
\begin{align}
  \BFcol{\mathrm{S}}{\mathrm{A}^\prime}(\hat{\theta}_r; g)
  &= \BFcol{\mathrm{S}}{\mathrm{A}}(\hat{\theta}_r; g)
  \frac{\Phi(\abs{z_o})}{\Phi\left\{\mathrm{sign}(z_o)
    %     \frac{z_o + z_r\sqrt{c}}{\sqrt{1 + c}} \right\}}
    \frac{z_o(1 + dc)}{\sqrt{1 + c}} \right\}}
   \label{eq1:BFrtrunc}
\end{align}
where $\Phi(\cdot)$ is the cumulative distribution function of the standard
normal distribution (see Appendix \ref{app:BFtrunc}). Hence, \eqref{eq1:BFrtrunc}
is the Bayes factor under the standard advocacy prior multiplied by a correction
term. Determining the smallest level of replication success
with~\eqref{eq1:BFrtrunc} leads to a corrected sceptical Bayes factor, while
setting $g = 0$ in~\eqref{eq1:BFrtrunc} leads to a corrected replication Bayes
factor. The correction term goes to one when the replication estimate goes in
the same direction as the original one and the replication sample size increases
($d > 0$ and $c \to \infty$), but it penalises when the replication estimate
goes in the opposite direction ($d < 0$ and $c \to \infty$).

@@

Figure~\ref{fig1:condpower} shows conditional and predictive power as a function
of the relative variance $c$ and for several values of of the original $z$-value
$z_o$ (respectively original minimum Bayes factor $\minBF_{o}$). In general,
uncertainty about replication success is higher for predictive power, leading it
to be closer to 50\% in all cases. As can also be seen, if the original result
was not convincing on its own (\eg if $z_o = 1.5$ or $z_{o} = 2$), it is
impossible to achieve replication success with the two-trials rule, the
sceptical Bayes factor, and the sceptical $p$-value. This is not the case for
the replication Bayes factor, for which high power can also be obtained for
small $z_o$ if $c$ is sufficiently large. However, as shown in the previous
section, the higher power of the replication Bayes factor comes at the cost of a
massive type I error inflation. For $z_o = 2.5$, the sceptical Bayes factor
shows higher power than the two-trials rule when $c = 1$, but the power of the
two-trials rule increases faster in $c$ and approaches the power curve of the
replication Bayes factor. The power of the sceptical $p$-value is still a bit
lower, likely due to the more stringent requirement on the minimum relative
effect estimate. For $z_{o} = 3$, the power differences between the methods
mostly disappear.

@@

\section{Extension to non-normal models} \label{sec1:tdist}
So far, we have always assumed approximate normality of the effect estimates
$\that_o$ and $\that_r$, as well as known variances $\sigma^2_o$ and
$\sigma^2_r$. This may be a problem for studies with small sample size and/or
extreme results (\eg when a study examines a rare disease with death rates close
to 0\%). One way of dealing with this issue is to consider the exact likelihood
of the data underlying the effect estimates, and then marginalise over possible
nuisance parameters \citep[Chapter 8.2.2]{Spiegelhalter2004}. This leads to
marginal likelihoods which are again only conditional on the effect size
$\theta$, allowing the procedures to be used analogously as described in the
proceeding sections. The choice of the likelihood depends on the type of effect
size $\theta$. We will illustrate the approach for \emph{standardised mean
differences} (SMD) and \emph{log odds ratios} (logOR), two of the most widely
used types of effect sizes.

>>>>

Let $\theta$ denote the effect of a treatment on an outcome of interest. Let
$\that_o$ and $\that_r$ denote its maximum likelihood estimates obtained from an
original (subscript $o$) and from a replication study (subscript $r$),
respectively. Let the corresponding standard errors be denoted by $\sigma_o$ and
$\sigma_r$, the $z$-values by $z_o = \that_o/\sigma_o$ and
$z_r = \that_r/\sigma_r$, and define the variance ratio as
$c = \sigma_o^2/\sigma_r^2$ and the relative effect estimate as
$d = \that_r/\that_o = z_r/(z_o\sqrt{c})$. For many effect size types the
variances are inversely proportional to the sample size, \ie{}
$\sigma^2_{o} = \kappa/n_{o}$ and $\sigma^2_{r} = \kappa/n_{r}$ for some unit
variance $\kappa$. The variance ratio is then the ratio of the replication to
the original sample size $c = n_{r}/n_{o}$.

We adopt a meta-analytic framework and consider the effect estimates as the
data, rather than their underlying samples, and assume that
$\that_k \given \theta \sim \Nor(\theta, \sigma^2_k)$ for $k \in \{o, r\},$ \ie{}
normality of the effect estimates around $\theta$, with known variances equal to
their squared standard errors. For studies with reasonable sample size, this
framework usually provides a good approximation for a wide range of (suitably
transformed) effect size types \citep[Chapter 2.4]{Spiegelhalter2004}. For
example, means and mean differences (no transformation), odds ratios, hazard
ratios, risk ratios (logarithmic transformation), or correlation coefficients
(``Fisher-$z$'' transformation). We refer to the literature of meta-analysis for
details about transformations of effect sizes \citep[\eg{}][Chapter
11.6]{Cooper2019}. The normal model in combination with conjugate priors enables
derivation of closed-form expressions in many cases, which allows us to easily
study limiting behaviour and facilitates interpretability. In Section
\ref{sec1:tdist}, we will present relaxations of the normality assumption, which
can lead to more accurate inferences when studies have small sample sizes and/or
show extreme results.

\section{Reverse-Bayes assessment of replication success with Bayes factors}
\label{sec1:methods}
% ``judgements are made concerning the values that the final
% probabilities would have after some imaginary results are obtained; then Bayes'
% theorem is used backwards to estimate what the consistent initial probabilities
% must be assumed to be; and lastly Bayes's theorem is used (forwards) to obtain
% estimates of the final probabilities in the light of the actual observations.
% This device takes \emph{seriously} the view that a theory of probability or of
% rationality is no more than a theory of consistency'' (p. 126).
The idea of reversing Bayes' theorem was first proposed by \citet{Good1950}. He
acknowledged that in many situations there is no obvious choice for the prior
distributions involved in Bayesian analyses. On the other hand, we are often
more certain which posterior inferences would convince us regarding the
credibility of a hypothesis. For this reason, Good inverted Bayes' theorem and
derived priors, which combined with the observed data, would lead to posterior
inferences that were specified beforehand (\eg{} the data favour one hypothesis
over another). His reverse-Bayes inference then centred around the question
whether the resulting prior is plausible, and if so, this would legitimise the
posterior inference. See Figure~\ref{fig1:RBschmema} for a graphical illustration
of this process.

@@

Good argued that philosophically there is nothing wrong with inferences
resulting from backwards use of Bayes' theorem, since the theorem merely
constrains prior and posterior to be consistent with the laws of probability
(regardless of their conventional names suggesting a particular temporal
ordering). Despite his advocacy, the reverse-Bayes idea remained largely
unexplored until \citet{Matthews2001b} introduced the \emph{Analysis of
Credibility}, which in turn led to new developments in reverse-Bayes methodology
(see \citet{Held2021b} for a recent review). Most of these approaches use the
reversal of Bayes' theorem in order to challenge or substantiate the credibility
of scientific claims. Usually, a posterior inference corresponding to
(non-)credibility of a claim is specified, and the associated prior is then
derived from the data. Inference is subsequently carried out based on this
reverse-Bayes prior, \eg{} the interest is often to check whether the prior is
plausible in light of external evidence, an obvious candidate being data from a
replication study. This can be done, for example, using methods to assess
prior-data conflict \citep{Box1980, Evans2006}.

@@

    \item \textbf{Replication study}:
          For the replication effect estimate
$\that_r \given \theta \sim \Nor(\theta, \sigma^2_r)$ compute the Bayes factor
contrasting the sceptic $\HS \colon \theta \sim \Nor(0, \gy \cdot \sigma^2_o)$
to an advocate $\HA \colon \theta \sim \Nor(\that_o, \sigma^2_o)$. Declare
\emph{replication success} at level $\gamma$ if
          $$\BFcol{\mathrm{S}}{\mathrm{A}}(\that_r; \gy) \leq \gamma,$$
          \ie{} the data favour the advocate over the sceptic at a higher level
than the sceptic's initial objection.

@@

The Bayes factor from equation ~\eqref{eq1:BFo} is shown in Figure ~\ref{fig1:bfo}
as a function of $g$ and for different original $z$-values $z_o$. For fixed
$z_o$, it is well known that this Bayes factor is bounded from below by
\begin{align}
  \label{eq1:minBFo}
  \minBF_{o} =
  \begin{cases}
    |z_o| \cdot  \exp(-z_o^2/2) \cdot \sqrt{e} & \text{for} ~ |z_o| > 1 \\
    1 & \text{for} ~ |z_o| \leq 1
  \end{cases}
\end{align}
which is reached at $\gminBFo = \max\{0, z^2_o - 1\}$ \citep{Edwards1963}.
Further increasing the relative sceptical prior variance increases
\eqref{eq1:BFo} indefinitely because of the Jeffreys-Lindley paradox, \ie{}
$\BFcol{0}{\mathrm{S}}(\hat{\theta}_o; g) \to \infty$ for $g \to \infty$
\citep[][Section 6.1.4]{Bernardo2000}. Hence, for a relative sceptical prior
variance $g \in [0, \gminBFo]$, the resulting Bayes factor will be
$\BFcol{0}{\mathrm{S}}(\hat{\theta}_o; g) \in [\minBF_o, 1]$.

@@

From Figure \ref{fig1:bfo} we can see that the more compelling the original data
(\ie{} the larger $|z_o|$), the smaller the sufficiently sceptical relative prior
variance $\gy$ needs to be in order to make the result no longer convincing at
level $\gamma$. In the most extreme case, when $|z_{o}| \to \infty$ and $\gamma$
remains fixed, the sufficiently sceptical prior variance will converge to
zero (Appendix \ref{appendix:lambertW}). On the other hand, if $|z_o|$ is not
sufficiently large, $\BFcol{0}{\mathrm{S}}(\hat{\theta}_o; g)$ will either be
always increasing in $g$ (if $|z_o| \leq 1$) or it will reach a minimum above
the chosen level $\gamma$. In both cases the sufficiently sceptical relative
prior variance $\gy$ is not defined since there is no need to challenge an
already unconvincing result.

A side note on the Jeffreys-Lindley paradox is worth being mentioned: If a
$\gy < \gminBFo$ exists, there exists also a $\gy^\prime > \gminBFo$ as the
Bayes factor monotonically increases in $g > \gminBFo$ and therefore must
intersect a second time with $\gamma$, due to the paradox. This means that the
more compelling the original result, the larger $\gy^\prime$ needs to be chosen,
such that the result becomes no longer convincing at level $\gamma$. However,
priors which become increasingly diffuse do not represent increasing scepticism
but rather increasing ignorance. Using \eqref{ggamma} therefore avoids this
manifestation of the Jeffreys-Lindley paradox, since it determines sceptical
priors only from the class of priors that become increasingly concentrated for
increasing evidence (\ie{} priors with $\gy \leq \gminBFo$). In principle, the
solution $\gy^\prime > \gminBFo $ could also be computed by replacing the
$\lw{-1}$ branch of the Lambert $W$ function in \eqref{ggamma} with the $\lw{0}$
branch, but this will not be of interest to us.

@@

For example, if we observe $z_o = 3$ (equivalent to minimum Bayes factor
$\minBF_{o} = 1/18$) and choose a level $\gamma = 1/10$
the sufficiently sceptical relative prior variance \eqref{ggamma} is
$\gy = 1.6$. If a replication is conducted with the
same precision ($c = 1$) and we observe $z_r = 2.5$ (equivalent to
minimum Bayes factor $\minBF_{r} = 1/5.5$ and relative effect
estimate $d = z_{r}/(z_{o}\sqrt{c}) = 0.83$), using equation
\eqref{eq1:BFr} this would lead to $\BFcol{\mathrm{S}}{\mathrm{A}}(\hat{\theta}_r;
1.6) = 1/3.5$, which means
that the replication was not successful at level $\gamma = 1/10$.
However, if we had chosen a less stringent level, \eg{}$\gamma =
1/3$, the replication would have been considered successful
since then $\gy = 0.4$ and
$\BFcol{\mathrm{S}}{\mathrm{A}}(\hat{\theta}_r; 0.4)
= 1/7.4$.

@@

\subsubsection{The two-trials rule}
Replication success is most commonly declared when both original and replication
study provide compelling evidence against a null effect. This approach is also
known as the \emph{two-trials rule} in drug development and usually a
requirement for drug approval \citep[Section 9.4]{Kay2015}. Most replication
projects report $p$-values associated with the effect estimates as measures of
evidence against the null, but also default Bayes factors have been used
\citep[see \eg{} the Bayesian supplement of][]{Camerer2018}. To compare the
two-trials rule with methods based on Bayes factors we will study the two-trials
rule based on the minimum Bayes factor from~\eqref{eq1:minBFo}, \ie{} replication
success at level $\gamma$ is established when both
$\minBF_{k} % = \abs{z_{k}}\exp(-z_{k}^{2}/2)\sqrt{e}
\leq \gamma$ for $k \in \{o, r\}$, as well as
$\sign(\hat{\theta}_o) = \sign(\hat{\theta}_r)$. This approach has a one-to-one
correspondence to the usual version of two-trials rule as minimum Bayes factors
and $p$-values both only depend on the $z$-values of original and replication
study.

The two-trials rule guarantees that both studies provide compelling evidence
against the null. Similarly, the sceptical Bayes factor requires the original
study to be compelling on its own since it can never be smaller than
$\minBF_{o}$. However, one can easily construct examples where the sceptical
Bayes factor is smaller than the minimum Bayes factor from the replication study
(\eg{} when $\minBF_{o} = 1/2$,
$\minBF_{r} = 1/1.5$, and $c = 1$
we obtain $\BFs = 1/1.9$). So for the same level of
replication success $\gamma$ the two-trials may not flag replication success
whereas the sceptical Bayes factor would.

@@

Similarly, as with the sceptical Bayes factor, the $Q$-statistic
from~\eqref{eq1:Qstat} appears in~\eqref{eq1:BFrep} and acts as a penalty term,
\ie{} larger values of $Q$ lower the degree of replication success. However, in
contrast to the sceptical Bayes factor, the replication Bayes factor is not
limited by the evidence from the original study because $\BFr \downarrow 0$ as
$c \to \infty$ provided $z_{o} \neq 0$ and $d \neq 0$. Moreover, we have that
\begin{align*}
  1 \geq \BFs \geq \BFcol{\mathrm{S}}{\mathrm{A}}(\hat{\theta}_r; g_{\scriptscriptstyle \BFs})
  = \BFcol{\mathrm{S}}{0}(\hat{\theta}_r; g_{\scriptscriptstyle \BFs}) \cdot \underbrace{\BFcol{0}{\mathrm{A}}(\hat{\theta}_r)}_{= \BFr}.
\end{align*}
So the sceptical Bayes factor is larger than the replication Bayes factor if the
replication data favour the sceptical prior
$\HS \colon \theta \sim \Nor(0, g_{\scriptscriptstyle \BFs} \cdot \sigma^{2}_{o})$
over the null hypothesis. They can only coincide when $\BFs = 1$ since then
$g_{\scriptscriptstyle \BFs} = 0$.

@@

\subsubsection{The sceptical \textit{p}-value}
Of particular interest is the relationship between the sceptical Bayes factor
and the sceptical $p$-value \citep{Held2020}, as it is the outcome of a similar
reverse-Bayes procedure. One also considers a sceptical prior for the effect
size $\theta \sim \Nor(0, \tau^2)$, the sufficiently sceptical prior variance at
level $\alpha$ is then defined as $\tau^2 = \tau^2_\alpha$ such that the
$(1 - \alpha)$ credible interval for $\theta$ based on the posterior
$\theta \given \that_o, \tau^2_\alpha$ does not include zero. Replication
success is declared if the tail probability of the replication effect estimate
under its prior predictive distribution
$\that_r \given \tau^2_\alpha \sim \Nor(0, \tau^2_\alpha + \sigma^2_r)$ is
smaller than $\alpha$. The smallest level $\alpha$ where replication success can
be established defines the sceptical $p$-value. In contrast to the sceptical
Bayes factor, the sceptical $p$-value always exists and there are closed form
expressions to compute it for all values of $c$, \ie{}
$\ps = 1 - \Phi(z_{\mathrm{S}})$ with
\begin{equation*}
  z^2_{\mathrm{S}} =
  \begin{cases}
    z^2_H/2 & \text{for}~c=1 \\ \frac{1}{c - 1}\left\{\sqrt{z^2_A\left[z^2_A +
          (c - 1)z^2_H\right]} - z^2_A\right\} & \text{for}~c \neq 1
  \end{cases}
  \label{eq1:z2s}
\end{equation*}
where $z^2_H = 2/(1/z_o^2 + 1/z_r^2)$ the harmonic mean,
$z^2_A = (z_o^2 + z_r^2)/2$ the arithmetic mean of the squared $z$-statistics,
and provided that $\sign(\that_o) = \sign(\that_r)$ (otherwise
$\ps = \Phi(z_{\mathrm{S}})$).

@@

In practice, the replication paradox is hardly an issue, since replications
rarely show such contradictory results, \eg{} to achieve replication success at
level $\gamma = 1/3$ with
$\minBF_{o} = 1/10$ and $c = 1$, the
relative effect estimate needs to be $d < -7.09$ for the
paradox to appear with the sceptical Bayes factor. The replication Bayes factor
is more prone to the paradox because its point-null hypothesis fails more
strongly to predict estimates in opposite direction, \eg{} for the same numbers as
before it requires $d < -2.66$.

In both cases the paradox can be overcome by truncating the advocacy prior $\HA$
such that only effects in the same direction as the original estimate $\that_o$
have non-zero probability, \ie{} for positive $\that_{o}$ consider
$\HAp \colon \theta \sim \Nor(\that_o, \sigma^2_o) \, \mathbb{1}_{(0, \infty)}(\theta)$,
where $\mathbb{1}_{B}(x)$ is the indicator function of the set $B$. The Bayes
factor contrasting $\HS$ to $\HAp$ turns out to be
\begin{align}
  \BFcol{\mathrm{S}}{\mathrm{A}^\prime}(\hat{\theta}_r; g)
  &= \BFcol{\mathrm{S}}{\mathrm{A}}(\hat{\theta}_r; g)
  \frac{\Phi(\abs{z_o})}{\Phi\left\{\mathrm{sign}(z_o)
    %     \frac{z_o + z_r\sqrt{c}}{\sqrt{1 + c}} \right\}}
    \frac{z_o(1 + dc)}{\sqrt{1 + c}} \right\}}
   \label{eq1:BFrtrunc}
\end{align}
where $\Phi(\cdot)$ is the cumulative distribution function of the standard
normal distribution (see Appendix \ref{app:BFtrunc}). Hence, \eqref{eq1:BFrtrunc}
is the Bayes factor under the standard advocacy prior multiplied by a correction
term. Determining the smallest level of replication success
with~\eqref{eq1:BFrtrunc} leads to a corrected sceptical Bayes factor, while
setting $g = 0$ in~\eqref{eq1:BFrtrunc} leads to a corrected replication Bayes
factor. The correction term goes to one when the replication estimate goes in
the same direction as the original one and the replication sample size increases
($d > 0$ and $c \to \infty$), but it penalises when the replication estimate
goes in the opposite direction ($d < 0$ and $c \to \infty$).

@@

Figure~\ref{fig1:condpower} shows conditional and predictive power as a function
of the relative variance $c$ and for several values of of the original $z$-value
$z_o$ (respectively original minimum Bayes factor $\minBF_{o}$). In general,
uncertainty about replication success is higher for predictive power, leading it
to be closer to 50\% in all cases. As can also be seen, if the original result
was not convincing on its own (\eg{} if $z_o = 1.5$ or $z_{o} = 2$), it is
impossible to achieve replication success with the two-trials rule, the
sceptical Bayes factor, and the sceptical $p$-value. This is not the case for
the replication Bayes factor, for which high power can also be obtained for
small $z_o$ if $c$ is sufficiently large. However, as shown in the previous
section, the higher power of the replication Bayes factor comes at the cost of a
massive type I error inflation. For $z_o = 2.5$, the sceptical Bayes factor
shows higher power than the two-trials rule when $c = 1$, but the power of the
two-trials rule increases faster in $c$ and approaches the power curve of the
replication Bayes factor. The power of the sceptical $p$-value is still a bit
lower, likely due to the more stringent requirement on the minimum relative
effect estimate. For $z_{o} = 3$, the power differences between the methods
mostly disappear.

@@

\section{Extension to non-normal models} \label{sec1:tdist}
So far, we have always assumed approximate normality of the effect estimates
$\that_o$ and $\that_r$, as well as known variances $\sigma^2_o$ and
$\sigma^2_r$. This may be a problem for studies with small sample size and/or
extreme results (\eg{} when a study examines a rare disease with death rates close
to 0\%). One way of dealing with this issue is to consider the exact likelihood
of the data underlying the effect estimates, and then marginalise over possible
nuisance parameters \citep[Chapter 8.2.2]{Spiegelhalter2004}. This leads to
marginal likelihoods which are again only conditional on the effect size
$\theta$, allowing the procedures to be used analogously as described in the
proceeding sections. The choice of the likelihood depends on the type of effect
size $\theta$. We will illustrate the approach for \emph{standardised mean
differences} (SMD) and \emph{log odds ratios} (logOR), two of the most widely
used types of effect sizes.
